---
title: Effect of climate warming on the timing of autumn leaf senescence reverses after the summer solstice
author: Constantin Zohner
date: "last updated March 31, 2023"

subtitle: Compare GPP/temperature effects with monthly water availability (Fig. S10)
output:
  html_document:
    highlight: haddock
    toc: false
    use_bookdown: true
    df_print: paged
---

<style>
body
  { counter-reset: source-line 0; }
pre.numberSource code
  { counter-reset: none; }
</style>

<br>

### Figure descriptions
- Fig. S10: The effects of pre-solstice GPP (upper panels) or temperature (lower panels) and monthly water availability (soil moisture in left panels; precipitation in right panels) on inter-annual variation in the timing of EOS10 from satellite observations. 

<br>

### Variable names
- Prcp...Precipitation
- Tday...Day-time temperature
- GPP...Gross primary productivity

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  class.source = "numberLines lineAnchors"
  )
```

<br><br>

```{r, message=FALSE, warning=FALSE, attr.source='.numberLines'}
###################
# Required packages
###################



require(tidyverse)
require(ggplot2)
require(data.table)
require(broom.mixed)
require(sjmisc)
require(lme4)



#define plot themes
plotTheme1 = theme(legend.position   = "none",
                   legend.background = element_rect(fill=NA, size=0.5, linetype="solid"),
                   legend.text       = element_text(color="black"),
                   panel.grid.major  = element_blank(),
                   panel.grid.minor  = element_blank(),
                   panel.background  = element_blank(),
                   panel.border      = element_rect(colour = "black", fill=NA),
                   axis.line         = element_line(color = "black"),
                   axis.text         = element_text(colour = "black"),
                   strip.background  = element_rect(fill=NA),
                   strip.text        = element_text(colour = 'black'),
                   plot.title        = element_text(face="bold"))



##############################################################################################################################################
##############################################################################################################################################



#####################
## Set directories ##
#####################



# set the working directory
setwd("/Users/consti/Desktop/PhD/Publication_material/17_Autumn_phenology_tier2/Remote_sensing/Analysis")

# paths
Drivers_path    = "Analysis_input/Drivers_final_EOS10/Merged_file"
output_path     = "Analysis_output_startSen/Monthly_water_availability"



##############################################################################################################################################
##############################################################################################################################################



#################
## Import data ##
#################



Pheno.df <- fread(paste(Drivers_path, "Remote_sensing_drivers_data_startSen_preseason.csv", sep="/")) 



##############################################################################################################################################
##############################################################################################################################################



##################################################################################################
## Test correlation between monthly (June-September) precipitation and pre-solstice temperature ##
##################################################################################################



ModelResults.df = Pheno.df %>%
  
  group_by(geometry) %>% 
  do({
    
    #run model
    ##########
    
    model1 = lm(Prcp6~Tday.LO.SO, data=.)
    model2 = lm(Prcp7~Tday.LO.SO, data=.)
    model3 = lm(Prcp8~Tday.LO.SO, data=.)
    model4 = lm(Prcp9~Tday.LO.SO, data=.)
    
    #create combined dataframe
    ##########################
    
    data.frame(rbind(
      
      glance(model1) %>% 
        mutate(var = 'June'),
      glance(model2) %>% 
        mutate(var = 'July'),
      glance(model3) %>% 
        mutate(var = 'August'),
      glance(model4) %>% 
        mutate(var = 'September')
    ) )
    
  }) 
  
ModelResults.df %>% 
  group_by(var) %>% 
  summarize(mean = round(mean(r.squared), 2))



##############################################################################################################################################
##############################################################################################################################################



##################
## Mixed models ##
##################



ModelResults.df = Pheno.df %>%
  do({
    
    #run model
    ##########
    
    model1 = lmer(scale(Senesc_DOY)~scale(Tday.LO.SO)+scale(Prcp6)+scale(Prcp7)+scale(Prcp8)+scale(Prcp9)+(1|geometry),  
                  data=., control = lmerControl(optimizer ="Nelder_Mead"))
    model2 = lmer(scale(Senesc_DOY)~scale(Tday.LO.SO)+scale(Moist6)+scale(Moist7)+scale(Moist8)+scale(Moist9)+(1|geometry),  
                  data=., control = lmerControl(optimizer ="Nelder_Mead"))
    model3 = lmer(scale(Senesc_DOY)~scale(GPP.LO.SO)+scale(Prcp6)+scale(Prcp7)+scale(Prcp8)+scale(Prcp9)+(1|geometry),  
                  data=., control = lmerControl(optimizer ="Nelder_Mead"))
    model4 = lmer(scale(Senesc_DOY)~scale(GPP.LO.SO)+scale(Moist6)+scale(Moist7)+scale(Moist8)+scale(Moist9)+(1|geometry),  
                  data=., control = lmerControl(optimizer ="Nelder_Mead"))
    
    #create combined dataframe
    ##########################
    
    data.frame(rbind(
      
      tidy(model1, effects="fixed") %>% 
        mutate(var1 = 'Precipitation',
               var2 = 'Temperature'),
      tidy(model2, effects="fixed") %>% 
        mutate(var1 = 'Moisture',
               var2 = 'Temperature'),
      tidy(model3, effects="fixed") %>% 
        mutate(var1 = 'Precipitation',
               var2 = 'GPP'),
      tidy(model4, effects="fixed") %>% 
        mutate(var1 = 'Moisture',
               var2 = 'GPP')
      ) )
    
  }) %>% 
  
  #delete intercept
  filter(!term %in% c("(Intercept)")) %>% 
  mutate(term = gsub("scale","",term),
         term = gsub("\\(|\\)","",term),
         term = plyr::revalue(term, c("Moist6"="June",
                                      "Moist7"='July',
                                      'Moist8'='August',
                                      'Moist9'='September',
                                      "Prcp6"="June",
                                      "Prcp7"='July',
                                      'Prcp8'='August',
                                      'Prcp9'='September',
                                      'GPP.LO.SO'='Pre-solstice',
                                      'Tday.LO.SO'='Pre-solstice')),
         term = factor(term, levels = c("Pre-solstice","June","July",'August',"September"))) 



##############################################################################################################################################
##############################################################################################################################################



##########
## Plot ##
##########


MonthlyWater.plot = ggplot(data = ModelResults.df, aes(x = term, y = estimate, fill=term)) + 
  geom_bar(stat = "identity")+
  geom_errorbar(aes(ymin = estimate - 2*std.error, ymax = estimate + 2*std.error), width=.2,
                position=position_dodge(.9)) +
  geom_hline(yintercept=0)+
  xlab("") + ylab("Standardized effect") +
  coord_cartesian(ylim = c(-0.42,0.42)) +
  scale_fill_manual(values = c('#F21A00',rep('#3B9AB2',4)))+
  plotTheme1 +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  facet_grid(var2~var1)


#save plots as .pdf
ggsave(MonthlyWater.plot , file="FigS10_Monthly_water.pdf", path=output_path,
       width=4.5, height=4.5)

MonthlyWater.plot



##############################################################################################################################################
##############################################################################################################################################



#####################
## Reproducibility ##	
#####################



## date time
Sys.time()

## session info
sessionInfo()



##############################################################################################################################################
#############################################################THE END##########################################################################
##############################################################################################################################################
```

